\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-09-04 paper.qmd
Updated Figure 14 (only figures)

trimesh_removed_pbmc <- ggplot() + 
  geom_segment(data = tr_from_to_df_pbmc, 
               aes(
                 x = x_from, 
                 y = y_from, 
                 xend = x_to, 
                 yend = y_to),
               colour = "#33a02c") +
  geom_point(
    data = umap_pbmc_scaled,
    aes(
      x = UMAP1,
      y = UMAP2
    ),
    alpha = 0.3,
    size = 0.1,
    color = "#000000"
  )  +
  coord_fixed() +
  interior_annotation("a")

@@

<!--bin1 = 30, bin2 = 29, b = 870, non_empty = 135-->
```{r}
#| echo: false
#| fig-cap: "(a) Model generated in $2\\text{-}D$ with UMAP, and (b) $p\\text{-}D$ model error in $2\\text{-}D$. The $2\\text{-}D$ model shows three well-separated distant clusters. The $p\\text{-}D$ model errors are distributed along clusters."
#| label: fig-model-pbmc-author
#| fig-pos: H
#| eval: false

trimesh_removed_pbmc + error_plot_pbmc + 
  plot_layout(guides='collect', ncol=2) &
  theme(legend.position='bottom')
```

@@

pbmc_proj_tsne_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.5,
    alpha = 0.2) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-1, 1)) +
  ylim(c(-1, 1)) +
  interior_annotation("b")

## Second projection
projection <- cbind(
    c(0.78689,-0.14774,0.36527,0.09617,0.12097,-0.00033,0.36973,-0.20959,-0.08438),
    c(0.18390,-0.43121,-0.30398,0.05861,0.55282,-0.14058,-0.27048,0.44386,-0.27372))

@@

pbmc_proj_tsne_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.5,
    alpha = 0.2) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-1, 1)) +
  ylim(c(-1, 1)) +
  interior_annotation("c")

## Third projection
projection <- cbind(
    c(0.00067,0.53963,-0.16241,-0.09777,0.32766,0.44543,-0.51359,-0.29884,0.00951),
    c(0.62399,0.26016,0.06441,-0.34686,-0.40466,0.24643,0.23537,0.07972,0.34402))

@@

pbmc_proj_tsne_model3 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.5,
    alpha = 0.2) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-1, 1)) +
  ylim(c(-1, 1)) +
  interior_annotation("d")

```{r}
#| label: fig-model-pbmc-author-proj
#| fig-pos: H
#| fig-height: 10
#| fig-width: 10
#| fig-cap: "Model in \\gD{}, on the UMAP layout (a), and three views of the fit in projections from $9\\text{-}D$ (b, c, d), for the PBMC3k data ($a_1 = 0.046$, $b = 676 \\  (26, \ 26)$, $m = 92$, benchmark value to remove large edges is $0.08$). The three well-separated clusters observed in the \\gD{} layout of UMAP (@fig-NLDR-variety (a)) is also visible when visualizing the model overlaid on the data space. But the separation between the clusters are small in $9\\text{-}D$. The model shows non-linear clusters and densed points within clusters in $9\\text{-}D$ space, which are some additional patterns not visible in the \\gD{} representation (@fig-NLDR-variety (a)). The full video is available at <https://youtu.be/VqqWuE0Jj6A>."

free(trimesh_removed_pbmc) + pbmc_proj_tsne_model1 +
  pbmc_proj_tsne_model2 + pbmc_proj_tsne_model3 +
  plot_layout(ncol = 2)
```

```{r}
#| label: tsne-pbmc
# read the png file from device

```{r}
#| label: tsne-pbmc-read
tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_30.rds")
pbmc_scaled_obj <- gen_scaled_data(
  data = tsne_pbmc)
tsne_pbmc_scaled <- pbmc_scaled_obj$scaled_nldr

<!--tSNE with perplexity:30-->
```{r}
#| echo: false
#| fig-cap: "The $2\\text{-}D$ layout from tSNE applied for the PBMC3k dataset. The default (hyper-)parameter setting is a perplexity $30$. We use our model-in-the-data-space to assess whether this is an accurate representation of structure present in the high-dimensional data, or if it is misleading."
#| label: fig-tsne-suggest
#| fig-pos: H
#| out-width: 40%
#| eval: false

tsne_pbmc 
```

pbmc_model <- fit_highd_model(
  training_data = training_data_pbmc,
  emb_df = tsne_pbmc_scaled,
  bin1 = num_bins_x_pbmc,
  r2 = r2_pbmc,
  q = 0.1,
  is_bin_centroid = TRUE,
  is_rm_lwd_hex = FALSE,
  col_start_highd = "PC"
)

benchmark_pbmc <- 0.1

tr_from_to_df_pbmc <- tr_from_to_df_pbmc |>
  filter(!(row_number() %in% c(155)))

trimesh_removed_pbmc <- ggplot() + 
  geom_segment(data = tr_from_to_df_pbmc, 
               aes(
                 x = x_from, 
                 y = y_from, 
                 xend = x_to, 
                 yend = y_to),
               colour = "#33a02c") +
  geom_point(
    data = tsne_pbmc_scaled,
    aes(
      x = tSNE1,
      y = tSNE2
    ),
    alpha = 0.2,
    color = "#000000"
  )  +
  coord_fixed() +
  interior_annotation("a")

## Compute error
error_df <- augment(
  df_bin_centroids = df_bin_centroids_pbmc,
  df_bin = df_bin_pbmc,
  training_data = training_data_pbmc,
  newdata = NULL,
  type_NLDR = "tSNE",
  col_start = "PC")

## To join embedding
error_df <- error_df |>
  bind_cols(tsne_pbmc_scaled |>
              select(-ID))

error_plot_pbmc <- error_df |>
  ggplot(aes(x = tSNE1,
             y = tSNE2,
             color = type,
             group = ID)) +
  geom_point(alpha=0.5,
             size = 0.1) +
  interior_annotation("b", sc_ltr_pos)

```{r}
#| label: pbmc-tsne-model-proj

## Removed long edge in high-d
distance_df_small_edges_pbmc <- distance_df_small_edges_pbmc |>
  filter(!(row_number() %in% c(155)))

## First projection
projection <- cbind(
    c(0.40456,-0.55243,-0.58434,0.05063,-0.20540,-0.10469,0.28850,0.12209,0.17017),
    c(0.34225,0.53615,-0.05822,-0.01993,-0.55480,0.21337,0.34100,-0.07280,-0.33139))

@@

pbmc_proj_umap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.5,
    alpha = 0.2) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-1, 1)) +
  ylim(c(-1, 1)) +
  interior_annotation("b")

## Second projection
projection <- cbind(
    c(0.07154,-0.44483,-0.04593,0.63343,0.21449,-0.39592,-0.29762,0.01726,0.30845),
    c(0.87788,-0.21837,0.11780,-0.17226,-0.17920,-0.00829,0.11424,0.28681,0.06091))

pbmc_proj_umap_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.5,
    alpha = 0.2) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-1, 1)) +
  ylim(c(-1, 1)) +
  interior_annotation("c")

## Third projection
projection <- cbind(
    c(0.13399,-0.00303,0.34841,-0.15626,-0.27432,0.48771,0.14252,0.39963,0.57983),
    c(0.49615,0.73920,-0.15299,-0.07463,-0.07948,0.13320,-0.15484,0.20042,-0.28869))

pbmc_proj_umap_model3 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.5,
    alpha = 0.2) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-1, 1)) +
  ylim(c(-1, 1)) +
  interior_annotation("d")

@@

```{r}
#| echo: false
#| fig-cap: "Model in \\gD{}, on the tSNE layout (a), and three views of the fit in projections from $9\\text{-}D$ (b, c, d), for the PBMC3k data ($a_1 = 0.099$, $b = 221 \\  (13, \ 17)$, $m = 92$, benchmark value to remove large edges is $0.1$). The three well-separated clusters with small separations observed in the \\gD{} layout of tSNE (@fig-NLDR-variety (e)) is also visible when visualizing the model overlaid on the data space in $9\\text{-}D$. Furthermore, there are some non-linear clusters. The model shows densed points within clusters in $9\\text{-}D$ space, which is an additional patterns not visible in the \\gD{} representation (@fig-NLDR-variety (e)). The full video is available at <https://youtu.be/5Y1hE4i7N2k>."
#| label: fig-pbmc2-sc
#| fig-pos: H
#| fig-width: 10
#| fig-height: 10

free(trimesh_removed_pbmc) + pbmc_proj_umap_model1 + 
  pbmc_proj_umap_model2 + pbmc_proj_umap_model3 +
  plot_layout(guides='collect', ncol=2) &
  theme(legend.position='none')
```

>>>>

trimesh_removed_pbmc <- ggplot() + 
  geom_segment(data = tr_from_to_df_pbmc, 
               aes(
                 x = x_from, 
                 y = y_from, 
                 xend = x_to, 
                 yend = y_to),
               colour = "#33a02c") +
  geom_point(
    data = umap_pbmc_scaled,
    aes(
      x = UMAP1,
      y = UMAP2
    ),
    alpha = 0.3,
    size = 0.1,
    color = "#000000"
  )  +
  coord_fixed() +
  interior_annotation("a1")

@@

@@

pbmc_proj_umap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.5,
    alpha = 0.2) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-1, 1)) +
  ylim(c(-1, 1)) +
  interior_annotation("a2")

## Second projection
projection <- cbind(
    c(0.00067,0.53963,-0.16241,-0.09777,0.32766,0.44543,-0.51359,-0.29884,0.00951),
    c(0.62399,0.26016,0.06441,-0.34686,-0.40466,0.24643,0.23537,0.07972,0.34402))

@@

pbmc_proj_umap_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.5,
    alpha = 0.2) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-1, 1)) +
  ylim(c(-1, 1)) +
  interior_annotation("a3")

```{r}
#| label: error-comp-pbmc
# read the png file from device

```{r}
#| label: umap-pbmc-read-best
umap_pbmc2 <- read_rds("data/pbmc3k/pbmc_umap_12_min_dist_0.99.rds")
pbmc_scaled_obj <- gen_scaled_data(
  data = umap_pbmc2)
umap_pbmc_scaled_best <- pbmc_scaled_obj$scaled_nldr

pbmc_model <- fit_highd_model(
  training_data = training_data_pbmc,
  emb_df = umap_pbmc_scaled_best,
  bin1 = num_bins_x_pbmc,
  r2 = r2_pbmc,
  q = 0.1,
  is_bin_centroid = TRUE,
  is_rm_lwd_hex = FALSE,
  col_start_highd = "PC"
)

#benchmark_pbmc <- 0.1

# tr_from_to_df_pbmc <- tr_from_to_df_pbmc |>
#   filter(!(row_number() %in% c(155)))

trimesh_removed_pbmc_best <- ggplot() + 
  geom_segment(data = tr_from_to_df_pbmc, 
               aes(
                 x = x_from, 
                 y = y_from, 
                 xend = x_to, 
                 yend = y_to),
               colour = "#33a02c") +
  geom_point(
    data = umap_pbmc_scaled_best,
    aes(
      x = UMAP1,
      y = UMAP2
    ),
    alpha = 0.2,
    color = "#000000"
  )  +
  coord_fixed() +
  interior_annotation("b1")

## Compute error
error_df <- augment(
  df_bin_centroids = df_bin_centroids_pbmc,
  df_bin = df_bin_pbmc,
  training_data = training_data_pbmc,
  newdata = NULL,
  type_NLDR = "UMAP",
  col_start = "PC")

## To join embedding
error_df <- error_df |>
  bind_cols(umap_pbmc_scaled_best |>
              select(-ID))

error_plot_pbmc_best <- error_df |>
  ggplot(aes(x = UMAP1,
             y = UMAP2,
             color = type,
             group = ID)) +
  geom_point(alpha=0.5,
             size = 0.1) +
  interior_annotation("b", sc_ltr_pos)

```{r}
#| label: pbmc-umap-model-proj-best

## Removed long edge in high-d
# distance_df_small_edges_pbmc <- distance_df_small_edges_pbmc |>
#   filter(!(row_number() %in% c(155)))

@@

pbmc_proj_umap_model1_best <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.5,
    alpha = 0.2) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-1, 1)) +
  ylim(c(-1, 1)) +
  interior_annotation("b2")

## Second projection
projection <- cbind(
    c(0.00067,0.53963,-0.16241,-0.09777,0.32766,0.44543,-0.51359,-0.29884,0.00951),
    c(0.62399,0.26016,0.06441,-0.34686,-0.40466,0.24643,0.23537,0.07972,0.34402))

@@

pbmc_proj_umap_model2_best <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.5,
    alpha = 0.2) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-1, 1)) +
  ylim(c(-1, 1)) +
  interior_annotation("b3")

@@

```{r}
#| echo: false
#| fig-cap: "Model in \\gD{}, on the tSNE layout (a), and three views of the fit in projections from $9\\text{-}D$ (b, c, d), for the PBMC3k data ($a_1 = 0.099$, $b = 221 \\  (13, \ 17)$, $m = 92$, benchmark value to remove large edges is $0.1$). The three well-separated clusters with small separations observed in the \\gD{} layout of tSNE (@fig-NLDR-variety (e)) is also visible when visualizing the model overlaid on the data space in $9\\text{-}D$. Furthermore, there are some non-linear clusters. The model shows densed points within clusters in $9\\text{-}D$ space, which is an additional patterns not visible in the \\gD{} representation (@fig-NLDR-variety (e)). The full video is available at <https://youtu.be/5Y1hE4i7N2k>."
#| label: fig-model-pbmc-author-proj
#| fig-pos: H
#| fig-width: 10
#| fig-height: 10

free(trimesh_removed_pbmc) + free(trimesh_removed_pbmc_best) +
  pbmc_proj_umap_model1 + pbmc_proj_umap_model1_best +
  pbmc_proj_umap_model2 + pbmc_proj_umap_model2_best +
  plot_layout(nrow=3) &
  theme(legend.position='none')
```

### 2024-09-04 paper.tex
Updated Figure 14 (only figures)

In order to find a reasonable NLDR representation for the PBMC3k
dataset, the MSE for different \(a_1\) using UMAP, tSNE, PHATE, PaCMAP,
and TriMAP with different (hyper-)parameter settings
(Figure~\ref{fig-pbmc-mse}) were calculated. tSNE with a perplexity
value set to \(30\), the default parameter setting, is considered as a
reasonable representation for the PBMC3k dataset. As shown in
Figure~\ref{fig-NLDR-variety} (e), there are three well-separated
clusters, although the separation between the clusters are small.
Additionally, non-linear structures can also be observed within the
clusters (Figure~\ref{fig-NLDR-variety} (e)). The visualization of the
\emph{model-in-the-data-space} provides evidence for this
(Figure~\ref{fig-pbmc2-sc}). Therefore, tSNE with perplexity \(30\)
accurately captures the data structure of the PBMC3k dataset, which is
better than UMAP.

\caption{\label{fig-model-pbmc-author-proj}Model in \gD{}, on the UMAP
layout (a), and three views of the fit in projections from
\(9\text{-}D\) (b, c, d), for the PBMC3k data (\(a_1 = 0.046\),
\(b = 676 \  (26,  26)\), \(m = 92\), benchmark value to remove large
edges is \(0.08\)). The three well-separated clusters observed in the
\gD{} layout of UMAP (Figure~\ref{fig-NLDR-variety} (a)) is also visible
when visualizing the model overlaid on the data space. But the
separation between the clusters are small in \(9\text{-}D\). The model
shows non-linear clusters and densed points within clusters in
\(9\text{-}D\) space, which are some additional patterns not visible in
the \gD{} representation (Figure~\ref{fig-NLDR-variety} (a)). The full
video is available at \url{https://youtu.be/VqqWuE0Jj6A}.}

@@

\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-pbmc2-sc-1.pdf}

@@

\caption{\label{fig-pbmc2-sc}Model in \gD{}, on the tSNE layout (a), and
three views of the fit in projections from \(9\text{-}D\) (b, c, d), for
the PBMC3k data (\(a_1 = 0.099\), \(b = 221 \  (13,  17)\), \(m = 92\),
benchmark value to remove large edges is \(0.1\)). The three
well-separated clusters with small separations observed in the \gD{}
layout of tSNE (Figure~\ref{fig-NLDR-variety} (e)) is also visible when
visualizing the model overlaid on the data space in \(9\text{-}D\).
Furthermore, there are some non-linear clusters. The model shows densed
points within clusters in \(9\text{-}D\) space, which is an additional
patterns not visible in the \gD{} representation
(Figure~\ref{fig-NLDR-variety} (e)). The full video is available at
\url{https://youtu.be/5Y1hE4i7N2k}.}

>>>>

In order to find a reasonable NLDR representation for the PBMC3k
dataset, the MSE for different \(a_1\) using UMAP, tSNE, PHATE, PaCMAP,
and TriMAP with different (hyper-)parameter settings
(Figure~\ref{fig-pbmc-mse}) were calculated. tSNE with a perplexity
value set to \(30\), the default parameter setting, is considered as a
reasonable representation for the PBMC3k dataset. As shown in
Figure~\ref{fig-NLDR-variety} (e), there are three well-separated
clusters, although the separation between the clusters are small.
Additionally, non-linear structures can also be observed within the
clusters (Figure~\ref{fig-NLDR-variety} (e)). The visualization of the
\emph{model-in-the-data-space} provides evidence for this
(\textbf{?@fig-pbmc2-sc}). Therefore, tSNE with perplexity \(30\)
accurately captures the data structure of the PBMC3k dataset, which is
better than UMAP.

@@

@@

\caption{\label{fig-model-pbmc-author-proj}Model in \gD{}, on the tSNE
layout (a), and three views of the fit in projections from
\(9\text{-}D\) (b, c, d), for the PBMC3k data (\(a_1 = 0.099\),
\(b = 221 \  (13,  17)\), \(m = 92\), benchmark value to remove large
edges is \(0.1\)). The three well-separated clusters with small
separations observed in the \gD{} layout of tSNE
(Figure~\ref{fig-NLDR-variety} (e)) is also visible when visualizing the
model overlaid on the data space in \(9\text{-}D\). Furthermore, there
are some non-linear clusters. The model shows densed points within
clusters in \(9\text{-}D\) space, which is an additional patterns not
visible in the \gD{} representation (Figure~\ref{fig-NLDR-variety} (e)).
The full video is available at \url{https://youtu.be/5Y1hE4i7N2k}.}
